---
title: ICC Initial Modeling Deck
author: "Shahryar Minhas"
date: "3/22/2018"
output:
  xaringan::moon_reader:
    css: "slides.css"
    lib_dir: libs
    nature:
      highlightStyle: github
      highlightLines: true
      countIncrementalSlides: false
---
exclude: true

```{r, message=FALSE, warning=FALSE, include=FALSE}
options(
  htmltools.dir.version = FALSE, # for blogdown
  width = 80,
  tibble.width = 80
)

knitr::opts_chunk$set(
  fig.align = "center"
)

library(emo)
htmltools::tagList(rmarkdown::html_dependency_font_awesome())
```

```{r setup, message=FALSE}
library(ggplot2)
theme_set(theme_bw())
library(dplyr)
library(stringr)
library(magrittr)
library(rvest)
```

---
class: middle
count: false

# Data

---

## DVs

- State
  + Prelim ICC Involvement: 18 countries experienced event between 2002 to 2016
  + Formal ICC Involvement: 5 countries experienced event between 2002 to 2016

- Opposition
  + Prelim ICC Involvement: 16 countries experienced event between 2002 to 2016
  + Formal ICC Involvement: 8 countries experienced event between 2002 to 2016

---

## IVs

- ICC ratification
- Africa region binary (static)
- GDP in constant dollars from World Bank (to 2017)
- Civil War & OSV from UCDP (to 2016)
- Political Terror Scale (to 2016)
- Polity (to 2016)
- Lower and higher court independence measures from VDEM (to 2017)
- Referral data (static)

---

## P5+1 Measures

For each P5+1 measure, I calculate the average relationship between every country in the sample and also include individual variables designating the relation between a country and each of the P5+1 countries.

- Ideal points from Voeten (to 2015)
- Military intervention from UCDP (to 2016)

Other affinity measures with P5+1 countries that we discussed include alliances, joint IGO membership, and trade. Each of these variables has too short a time series. I have the SIPRI measure set up as well but am hesitant about using it for now because the level of missingness in the dataset exceeds 30%.

---

## Effective sample

- Lagging each of the IVs and P5+1 measures by one year provides us with an effective sample from 2002 to 2016. 
- Each of the control and key variables of interest come with degrees of missingness, but in no case do we have missingness for more than 10% of cases. 
- However, given that some of those missing cases will lead us to lose the already number of rare events that we have I imputed five missing datasets using a Bayesian copula scheme. 

---

# Modeling

---

## Strategies

- We've discussed a number of modeling strategies:
  + Logit models estimated separately for prelim and formal stage
  + Logit model with temporal dependence taken into account using cubic splines (Carter & Signorino, 2010)
  + Selection model to account for the conditional dependence between the formal and prelim stages
  + Matching strategies given the covariates we assembled
  + Random forest models to aid in accounting for paucity of events

## Aside on selection models

I tried setting up a Heckman selection model with varying covariates in the prelim and formal stages but the correlation across the equations tended to be above 0.8 no matter the specification. I switched to a Sartori model, which relaxes the constraint that the 

For now I've worked on using a Sartori's version of a selection model since we haven't discussed whether we have enough covariates to specify separate theories for the prelim and formal equations

---

## Aside on matching



---

## Model specification

In general, the model specification I used for each of the above is: 

---