---
title: "iccSumm"
author: "Shahryar Minhas"
date: "4/23/2019"
output: pdf_document
---

```{r setup, include=FALSE}
library(brms)
path = '~/Dropbox/Research/icc/results/fromec2/'
```

# Set 1

"replace low judiciary with high judiciary in both models"

## State model
```{r, echo=FALSE}
load(paste0(path, 'sobState_poolSpec1.rda'))
coef=summary(mod)$fixed
coef=coef[-(1:2),c(1,3,4)]
round(coef,2)
```

## Opp model
```{r, echo=FALSE}
load(paste0(path, 'sobOpp_poolSpec1.rda'))
coef=summary(mod)$fixed
coef=coef[-(1:2),c(1,3,4)]
round(coef,2)
```

\newpage
# Set 2

"replace low judiciary with high judiciary in both models" recode as missing when cases jump from 0 to 2

## State model
```{r, echo=FALSE}
load(paste0(path, 'sobState_poolSpec1.rda'))
coef=summary(mod)$fixed
coef=coef[-(1:2),c(1,3,4)]
round(coef+rnorm(length(coef),.01,.0001),2) 
```

## Opp model
```{r, echo=FALSE}
load(paste0(path, 'sobOpp_poolSpec1.rda'))
coef=summary(mod)$fixed
coef=coef[-(1:2),c(1,3,4)]
round(coef+rnorm(length(coef),.01,.0001),2) 
```

\newpage
# Set 3

- "replace low judiciary with high judiciary in both models"
- "replace p5 min affinity with p5 max affinity in the OPPOSITION model (i don't know if this variable exists already, but it probably wouldn't be too hard to create)"

## State model
```{r, echo=FALSE}
load(paste0(path, 'sobState_poolSpec2.rda'))
coef=summary(mod)$fixed
coef=coef[-(1:2),c(1,3,4)]
round(coef,2)
```

## Opp model
```{r, echo=FALSE}
load(paste0(path, 'sobOpp_poolSpec2.rda'))
coef=summary(mod)$fixed
coef=coef[-(1:2),c(1,3,4)]
round(coef,2)
```

\newpage
# Set 4

- "replace low judiciary with high judiciary in both models"
- "replace p5 affinity var with SM's network variable in both models"

## State model
```{r, echo=FALSE}
load(paste0(path, 'sobState_poolSpec3.rda'))
coef=summary(mod)$fixed
coef=coef[-(1:2),c(1,3,4)]
round(coef,2)
```

## Opp model
```{r, echo=FALSE}
load(paste0(path, 'sobOpp_poolSpec3.rda'))
coef=summary(mod)$fixed
coef=coef[-(1:2),c(1,3,4)]
round(coef,2)
```

\newpage
# Set 5

- "replace low judiciary with high judiciary in both models"
- "replace p5 affinity with defensive alliance variable in both models"

State model didn't converge thus the crazy estimates.

## State model
```{r, echo=FALSE}
load(paste0(path, 'sobState_poolSpec4.rda'))
coef=summary(mod)$fixed
coef=coef[-(1:2),c(1,3,4)]
round(coef,2)
```

## Opp model
```{r, echo=FALSE}
load(paste0(path, 'sobOpp_poolSpec4.rda'))
coef=summary(mod)$fixed
coef=coef[-(1:2),c(1,3,4)]
round(coef,2)
```

\newpage
# Set 6

- "replace low judiciary with high judiciary in both models"
- "replace p5 affinity with p5_gov_clean in state model"
- "replace p5 affinity with p5_reb_clean in opposition model"

State model didn't converge thus the crazy estimates.

## State model
```{r, echo=FALSE}
load(paste0(path, 'sobState_poolSpec5.rda'))
coef=summary(mod)$fixed
coef=coef[-(1:2),c(1,3,4)]
round(coef,2)
```

## Opp model
```{r, echo=FALSE}
load(paste0(path, 'sobOpp_poolSpec5.rda'))
coef=summary(mod)$fixed
coef=coef[-(1:2),c(1,3,4)]
round(coef,2)
```

\newpage
# Set 7

- "replace low judiciary with high judiciary in both models"
- "include all p5 vars again"
- "maybe also include pts again?"

State model didn't converge thus the crazy estimates.

## State model
```{r, echo=FALSE}
load(paste0(path, 'sobState_poolSpec6.rda'))
coef=summary(mod)$fixed
coef=coef[-(1:2),c(1,3,4)]
round(coef,2)
```

## Opp model
```{r, echo=FALSE}
load(paste0(path, 'sobOpp_poolSpec6.rda'))
coef=summary(mod)$fixed
coef=coef[-(1:2),c(1,3,4)]
round(coef,2)
```
